---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading required libraries

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)

# disable messages
knitr::opts_chunk$set(message = FALSE)
```

## Loading and preprocessing the data

```{r load_data}
unzip("activity.zip", exdir = "data")
activity <- read.csv("data/activity.csv") %>% as_tibble()
activity <- activity %>% mutate(date = ymd(date))
head(activity)
```
The dates are parsed according to YYYY-MM-DD format.

## What is mean total number of steps taken per day?

```{r daily_steps}
# get total steps per day
steps_by_day <- activity %>% 
    group_by(date) %>% 
    summarise(total_steps = sum(steps, na.rm = TRUE))

mean_daily_steps <- mean(steps_by_day$total_steps)
median_daily_steps <- median(steps_by_day$total_steps)

steps_by_day %>% ggplot(aes(x = total_steps)) +
    geom_histogram(binwidth = 1000) +
    labs(x = "Total steps", y = "Count",
         title = "Distribution of total steps per day")
```

**Mean** total number of steps taken per day = **`r mean_daily_steps`** \
**Median** total number of steps taken per day = **`r median_daily_steps`**

## What is the average daily activity pattern?

```{r interval_steps}
# get the average number of steps per interval, across all days
steps_by_interval <- activity %>% 
    group_by(interval) %>% 
    summarise(avg_steps = mean(steps, na.rm = TRUE))

max_steps_interval <- steps_by_interval %>%
    filter(avg_steps == max(avg_steps)) %>% 
    pull(interval)

steps_by_interval %>% ggplot(aes(x = interval, y = avg_steps)) +
    geom_line() +
    labs(x = "Interval", y = "Average steps",
         title = "Number of steps in each interval, averaged across days")
```

On average, the interval **`r max_steps_interval`** contains the **maximum**
number of steps.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
